A GRASP algorithm for the concrete delivery problem

### 2024-04-16 ArticleCDP/journals/elsevier/articleCDP.tex
Article soumis

Academic research on concrete batch and delivery began in the late 1990s. \cite{tommelein1999just} described RMC as a prototypical example of a JIT production system in construction and identified two practices for delivering it. One approach is for the customer to haul the product from the batch plant with their concrete mixer, while the other is for the batch plant to deliver the concrete directly to the customer's location. This latter approach is the one that has been studied in all related papers found in the literature. Several works to schedule and dispatch concrete production and delivery have mainly focused on simulation-related methods. These methods can be standalone, such as those used by \cite{zayed2001simulation} and \cite{wang2001scheduling}, among others. Alternatively, these methods can be hybridized with optimization techniques, such as those used by \cite{feng2004optimizing} and \cite{lu2005optimized}. \cite{wang2001scheduling} developed a simulation model to reveal the effect and value of the concrete mixers' inter-arrival time on the productivity of hired unloading equipment on site. \cite{feng2004optimizing} used a combination of genetic algorithm (GA) and simulation process to minimize the total waiting time for trucks at a customer site. The study focused on loading homogeneous trucks at the same batch plant, with fixed loading and unloading durations. The GA was used to find the best loading sequence of RMC trucks to be assigned to different construction sites. The simulation process determined the loading, arrival, departure, and waiting time of trucks and thus evaluated the cost of each dispatching sequence. They evaluated their method using data from a batch plant in Taiwan with up to nine customers served. \cite{lu2005optimized} used the same combination of GA and simulation to determine the optimal number of concrete mixers to be deployed and an optimal schedule for batching and delivering concrete. Their objective was to minimize the idle time of the site crew due to late concrete deliveries and truck queuing time. In this setting, it was also necessary to deliver a batch of mortar on-site to lubricate the unloading pump before the concrete delivery. As such, the simulation model also included the batch and delivery of mortar.

@@

Let $n_o$ be the number of deliveries needed to fulfill the order $o$. $n_o$ is not known in advance because the trucks have different capacities. However, we can compute lower $n_o^{min}$ and upper $n_o^{max}$ bounds using the capacities of the largest $Q_{max}$ and smallest $Q_{min}$ available trucks.
\begin{alignat}{3}
    \label{mod:c0}
    n_o^{min} = \left\lceil \frac{q_o}{Q_{max}} \right\rceil \leq n_o \leq n_o^{max} = \left\lceil \frac{q_o}{Q_{min}} \right\rceil & \text{ } &
    \forall  o \in O.
\end{alignat}

@@

    \label{mod:c1}         & \sum_{j \in \delta^{+}(i) }{x^k_{ij}} =1   & \hspace{-10mm}  \forall i \in S, k \in K   \\
    \label{mod:c2}         & \sum_{j \in \delta^{-}(i)}{x^k_{ji}} = 1     & \hspace{-10mm} \forall i \in E, k \in K\\

@@

    \label{mod:c3}         & v_j \geq  w_i + \alpha_{b} + t_{ij} - M\left(1-x^{k}_{ij}\right) & \hspace{-40mm} \forall b \in B, i \in L_b, j \in \delta^{+}(i),  k \in K     \\
    \label{mod:c4}         & v_j \geq  w_i + \rho + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                       & \hspace{-40mm} \forall i \in D  , j \in \delta^{+}(i), k \in K                         \\
    \label{mod:c5}         & w_{i} \geq v_{i}  + \frac{q^k_j}{\tau^l_b} -M\left(1- x^{k}_{ij}\right)   & \hspace{-40mm} \forall  b \in B,  i \in L_{b},  j \in \delta^{+}(i), k \in K \\
    \label{mod:c6}         & w_{j} \geq v_{j}  + {  \frac{q^k_j}{\tau^u_c} -M\left(1- x^{k}_{ij}\right) }     & \hspace{-40mm}  \forall c \in  C, j \in D_{c}, i \in \delta^{-}(j),  k \in K \\

@@

    \label{mod:c8}         & v_{d^0_{o}} \geq a_c    & \hspace{-50mm} \forall  c \in C, \forall o \in O_c \text{ }                           \\
    \label{mod:c9}          & g_{c} \geq v_{d^0_{o_1}} - a_c - M\left(  \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})  }{u_{jd^0_{o_1}}}\right)  & \hspace{-50mm}  \forall  c \in C, \forall o_1 \in O_c \text{ }  \\
    \label{mod:c91}  & g_{c} - a_c \leq \tau^w_c  & \hspace{-20mm}  \forall  c \in C \text{ }   \\
    \label{mod:c101}       & v_{d^0_{o_1}} \geq w_j - M\left(1- u_{jd^0_{o_1}}\right)   & \hspace{-50mm}    \forall o_1 \in O_i,  j \in \delta_{D}^{-}(d^0_{o_1}) \text{ }                \\
    \label{mod:c10}        & v_{d^0_{o_1}}  \leq  w_j + \gamma^2 + M\left(1- u_{j,d^0_{o_1}}\right)     & \hspace{-40mm}  \forall  o_1 \in O,  j \in \delta_{D}^{-}(d^0_{o_1}) \text{ }                    \\
    \label{mod:c11}        & \sum_{ \substack{o_1 \in O_c }}{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } = |O_c|-1      & \hspace{-40mm} \forall  c \in C,|O_c| > 1   \text{ }   \\
    \label{mod:c12}        & \sum_{ \substack{o_1 \in O_c }}\sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}{u_{d^0_{o_1},j } } = |O_c|-1      & \hspace{-40mm} \forall  c \in C, |O_c| > 1  \text{ }          \\
    \label{mod:c13}    & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } \leq 1    & \hspace{-20mm}  \forall o_1 \in O \text{ }         \\
    \label{mod:c14}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \leq 1      & \hspace{-20mm} \forall o_1 \in O \text{ }        \\
    \label{mod:c15}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } +\smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \geq 1    & \hspace{-50mm} \forall o_1 \in O \text{ }    \\
    \label{mod:c16}        & v_{j+1} \geq w_{j} - M\left(1- u_{j,j+1}\right)   & \hspace{-50mm}  \forall o \in O, j \in D_{o}, j \geq 1  \text{ }    \\
    \label{mod:c17}        & v_{j+1}  \leq  w_{j} + \gamma^1 + M\left(1- u_{j,j+1}\right)    & \hspace{-60mm} \forall c \in C, \forall o \in O_c, j \in D_{o}, j \geq 1 \text{ }     \\
    \label{mod:c18}        & u_{j,j+1} \geq u_{j+1,j+2}      & \hspace{-70mm} \forall o_1 \in O,  j \in D_{o_1}, 1 \leq j \leq n_{o_{1}}-2 \text{ }   \\
    \label{mod:c19}        & u_{j,j+1} \geq \sum_{l \in L}{x_{lj}}   & \hspace{-60mm}  \forall o_1 \in O, j \in D_{o_1}, j \geq 1 \text{ }                   \\
    \label{mod:c20}        & v_{j} \geq w_{i} - M\left(1- u_{i,j}\right)   & \hspace{-40mm}  \forall i \in L, j \in \delta^{+}_{L}(i) \text{ }  \\
    \label{mod:c21}        & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \leq 1     & \hspace{-40mm}  \forall i \in L \text{ }         \\
    \label{mod:c22}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \leq 1                                      & \hspace{-40mm} \forall i \in L           \text{ }       \\
    \label{mod:c230}       & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } }  \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}}    & \hspace{-40mm} \forall i \in L \text{ }                                                      \\
    \label{mod:c231}       & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \geq \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i+1)}}{u_{j,i+1 } }   & \hspace{-20mm} \forall i \in L  \text{ }                                                        \\
    \label{mod:c23}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } + \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}}    & \hspace{-20mm} \forall i \in L  \text{ }                                                        \\
    \label{mod:c24}        & \sum_{k \in K}\sum_{j \in D_o}{q^k_j} = q_o            & \hspace{-40mm}  \forall o \in O  \text{ }           \\
    \label{mod:c240}       & q^k_j \leq \sum_{i \in  \delta^{-}(j)}{Q^k x^k_{ij}}   & \hspace{-40mm}  \forall j \in D, k \in K  \text{ }                                              \\

@@

    \label{mod:c29}        & w^1_k \geq  \mu + h_k - v_{e_k}            & \hspace{-40mm} \forall k \in K   \text{ }         \\
    \label{mod:c30}        & w^2_k \geq (v_{e_k} - h_k) - \eta   & \hspace{-40mm} \forall k \in K         \text{ }         \\

@@

            $Sol \leftarrow Sol \backslash \{(L^k_{i}, U^k_{i}), \forall i \in D_c\}$

@@

$Sol \leftarrow Sol \backslash \{(L^k_{i}, U^k_{i}), \forall i \in (D^{j'}_{o},..,D^{j-1}_{o})  \}$

@@

            $Sol \leftarrow Sol \backslash \{(L^k_j,U^k_j), \forall j \in D_c \}$
           }
           $CL \leftarrow CL\backslash CL'$ 
        }
        \uElse{
            $c_{min} \leftarrow \min\left\{ cost \ | \ (L^k_{d}, U^k_{d},cost,k) \in T   \right\} $, $c_{max} \leftarrow \max\left\{ cost \ | \ (L^k_{d}, U^k_{d},cost,k) \in T   \right\} $

@@

We also implemented two moves based on the relocate operator. The Consolidate Load operator is applied on an order with a remaining demand between $]Q_{min},Q_{max}]$. If this remaining demand is served with two drivers with capacity $Q_{min}$, then the operator tries to consolidate it to be served by a single driver with a capacity of $Q_{max}$. Next, the Relocate Driver operator removes a delivery node from its current route and inserts it in another driver's route. This enables the redistribution of delivery nodes among drivers, potentially reducing drivers underutilization and overtime costs.
% \clearpage
\section{Computational experiments}
\label{sec:cdp_comp_exp}

>>>>

Academic research on concrete batch and delivery began in the late 1990s. \cite{tommelein1999just} described RMC as a prototypical example of a JIT production system in construction and identified two practices for delivering it. One approach is for the customer to haul the product from the batch plant with their concrete mixer, while the other is for the batch plant to deliver the concrete directly to the customer's location. This latter approach is the one that has been studied in all related papers found in the literature. Several works to schedule and dispatch concrete production and delivery have mainly focused on simulation-related methods. These methods can be standalone \citep{zayed2001simulation} or can be hybridized with optimization techniques, such as those used by \cite{feng2004optimizing} and \cite{lu2005optimized}. \cite{feng2004optimizing} used a combination of genetic algorithm (GA) and simulation process to minimize the total waiting time for trucks at a customer site. The study focused on loading homogeneous trucks at the same batch plant, with fixed loading and unloading durations. The GA was used to find the best loading sequence of RMC trucks to be assigned to different construction sites. The simulation process determined the loading, arrival, departure, and waiting time of trucks and thus evaluated the cost of each dispatching sequence. They evaluated their method using data from a batch plant in Taiwan with up to nine customers served. \cite{lu2005optimized} used the same combination of GA and simulation to determine the optimal number of concrete mixers to be deployed and an optimal schedule for batching and delivering concrete. Their objective was to minimize the idle time of the site crew due to late concrete deliveries and truck queuing time. In this setting, it was also necessary to deliver a batch of mortar on-site to lubricate the unloading pump before the concrete delivery. As such, the simulation model also included the batch and delivery of mortar.

@@

Let $n_o$ be the number of deliveries needed to fulfill the order $o$. $n_o$ is not known in advance because the trucks have different capacities. However, we can compute lower $n_o^{min}$ and upper $n_o^{max}$ bounds using the capacities of the largest $Q_{max}$ and smallest $Q_{min}$ available trucks.
\begin{alignat}{3}
    \label{mod:c0}
    n_o^{min} = \left\lceil \frac{q_o}{Q_{max}} \right\rceil \leq n_o \leq n_o^{max} = \left\lceil \frac{q_o}{Q_{min}} \right\rceil.
\end{alignat}

@@

    \label{mod:c1}         & \sum_{j \in \delta^{+}(i) }{x^k_{ij}} =1   & \hspace{-10mm}  i \in S, k \in K   \\
    \label{mod:c2}         & \sum_{j \in \delta^{-}(i)}{x^k_{ji}} = 1     & \hspace{-10mm} i \in E, k \in K\\

@@

    \label{mod:c3}         & v_j \geq  w_i + \alpha_{b} + t_{ij} - M\left(1-x^{k}_{ij}\right) & \hspace{-40mm} b \in B, i \in L_b, j \in \delta^{+}(i),  k \in K     \\
    \label{mod:c4}         & v_j \geq  w_i + \rho + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                       & \hspace{-40mm} i \in D  , j \in \delta^{+}(i), k \in K                         \\
    \label{mod:c5}         & w_{i} \geq v_{i}  + \frac{q^k_j}{\tau^l_b} -M\left(1- x^{k}_{ij}\right)   & \hspace{-40mm}  b \in B,  i \in L_{b},  j \in \delta^{+}(i), k \in K \\
    \label{mod:c6}         & w_{j} \geq v_{j}  + {  \frac{q^k_j}{\tau^u_c} -M\left(1- x^{k}_{ij}\right) }     & \hspace{-40mm}  c \in  C, j \in D_{c}, i \in \delta^{-}(j),  k \in K \\

@@

    \label{mod:c8}         & v_{d^0_{o}} \geq a_c    & \hspace{-50mm}  c \in C, o \in O_c \text{ }                           \\
    \label{mod:c9}          & g_{c} \geq v_{d^0_{o_1}} - a_c - M\left(  \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})  }{u_{jd^0_{o_1}}}\right)  & \hspace{-50mm}   c \in C, o_1 \in O_c \text{ }  \\
    \label{mod:c91}  & g_{c} - a_c \leq \tau^w_c  & \hspace{-20mm}   c \in C \text{ }   \\
    \label{mod:c101}       & v_{d^0_{o_1}} \geq w_j - M\left(1- u_{jd^0_{o_1}}\right)   & \hspace{-50mm}    o_1 \in O_i,  j \in \delta_{D}^{-}(d^0_{o_1}) \text{ }                \\
    \label{mod:c10}        & v_{d^0_{o_1}}  \leq  w_j + \gamma^2 + M\left(1- u_{j,d^0_{o_1}}\right)     & \hspace{-40mm}   o_1 \in O,  j \in \delta_{D}^{-}(d^0_{o_1}) \text{ }                    \\
    \label{mod:c11}        & \sum_{ \substack{o_1 \in O_c }}{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } = |O_c|-1      & \hspace{-40mm}  c \in C,|O_c| > 1   \text{ }   \\
    \label{mod:c12}        & \sum_{ \substack{o_1 \in O_c }}\sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}{u_{d^0_{o_1},j } } = |O_c|-1      & \hspace{-40mm}  c \in C, |O_c| > 1  \text{ }          \\
    \label{mod:c13}    & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } \leq 1    & \hspace{-20mm}  o_1 \in O \text{ }         \\
    \label{mod:c14}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \leq 1      & \hspace{-20mm} o_1 \in O \text{ }        \\
    \label{mod:c15}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } +\smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \geq 1    & \hspace{-50mm} o_1 \in O \text{ }    \\
    \label{mod:c16}        & v_{j+1} \geq w_{j} - M\left(1- u_{j,j+1}\right)   & \hspace{-50mm}  o \in O, j \in D_{o}, j \geq 1  \text{ }    \\
    \label{mod:c17}        & v_{j+1}  \leq  w_{j} + \gamma^1 + M\left(1- u_{j,j+1}\right)    & \hspace{-60mm} c \in C, o \in O_c, j \in D_{o}, j \geq 1 \text{ }     \\
    \label{mod:c18}        & u_{j,j+1} \geq u_{j+1,j+2}      & \hspace{-70mm} o_1 \in O,  j \in D_{o_1}, 1 \leq j \leq n_{o_{1}}-2 \text{ }   \\
    \label{mod:c19}        & u_{j,j+1} \geq \sum_{l \in L}{x_{lj}}   & \hspace{-60mm}  o_1 \in O, j \in D_{o_1}, j \geq 1 \text{ }                   \\
    \label{mod:c20}        & v_{j} \geq w_{i} - M\left(1- u_{i,j}\right)   & \hspace{-40mm}  i \in L, j \in \delta^{+}_{L}(i) \text{ }  \\
    \label{mod:c21}        & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \leq 1     & \hspace{-40mm}  i \in L \text{ }         \\
    \label{mod:c22}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \leq 1                                      & \hspace{-40mm} i \in L           \text{ }       \\
    \label{mod:c230}       & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } }  \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}}    & \hspace{-40mm} i \in L \text{ }                                                      \\
    \label{mod:c231}       & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \geq \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i+1)}}{u_{j,i+1 } }   & \hspace{-20mm} i \in L  \text{ }                                                        \\
    \label{mod:c23}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } + \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}}    & \hspace{-20mm} i \in L  \text{ }                                                        \\
    \label{mod:c24}        & \sum_{k \in K}\sum_{j \in D_o}{q^k_j} = q_o            & \hspace{-40mm}  o \in O  \text{ }           \\
    \label{mod:c240}       & q^k_j \leq \sum_{i \in  \delta^{-}(j)}{Q^k x^k_{ij}}   & \hspace{-40mm}  j \in D, k \in K  \text{ }                                              \\

@@

    \label{mod:c29}        & w^1_k \geq  \mu + h_k - v_{e_k}            & \hspace{-40mm} k \in K   \text{ }         \\
    \label{mod:c30}        & w^2_k \geq (v_{e_k} - h_k) - \eta   & \hspace{-40mm} k \in K         \text{ }         \\

@@

            $Sol \leftarrow Sol \backslash \{(L^k_{i}, U^k_{i}), i \in D_c\}$

@@

$Sol \leftarrow Sol \backslash \{(L^k_{i}, U^k_{i}), i \in (D^{j'}_{o},..,D^{j-1}_{o})  \}$

@@

            $Sol \leftarrow Sol \backslash \{(L^k_j,U^k_j), j \in D_c \}$
           }
           $CL \leftarrow CL\backslash CL'$
        }
        \uElse{
            $c_{min} \leftarrow \min\left\{ cost \ | \ (L^k_{d}, U^k_{d},cost,k) \in T   \right\} $, $c_{max} \leftarrow \max\left\{ cost \ | \ (L^k_{d}, U^k_{d},cost,k) \in T   \right\} $

@@

We also implemented two moves based on the relocate operator. The Consolidate Load operator is applied on an order with a remaining demand between $]Q_{min},Q_{max}]$. If this remaining demand is served with two drivers with capacity $Q_{min}$, then the operator tries to consolidate it to be served by a single driver with a capacity of $Q_{max}$. Next, the Relocate Driver operator removes a delivery node from its current route and inserts it in another driver's route. This enables the redistribution of delivery nodes among drivers, potentially reducing drivers underutilization and overtime costs.
% \clearpage

\section{Computational experiments}
\label{sec:cdp_comp_exp}
